---
title: "HW 6"
author: "Christina Dai"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

# Problem 2

###Importing weather data:

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

### Bootstrapping 5000 samples from the dataset and producting estimates of r^2 and log(beta1 * beta2):

```{r}
weather_boot_results = 
  weather_df %>% 
    bootstrap(1000, id = "strap_number") %>% 
    mutate(
      models = map(strap, ~lm(tmax ~ tmin + prcp, data = .x)),
      results_r2 = map(models, broom::glance),
      results_logb1b2 = map(models, broom::tidy)
    ) %>% 
    unnest(results_r2) %>% 
    janitor::clean_names() %>% 
    select(strap_number, r_squared, results_logb1b2) %>% 
    unnest(results_logb1b2) %>% 
    janitor::clean_names() %>% 
    select(strap_number:estimate) %>% 
    pivot_wider(
      names_from = term,
      values_from = estimate
    ) %>% 
    mutate(log_b1b2 = log(tmin * prcp)) %>% 
    select(-`(Intercept)`)
```


### Plotting the estimates:

R_squared distribution

```{r}
weather_boot_results %>% 
  ggplot(aes(x = r_squared)) + 
  geom_density() +
  labs(title = "R_squared distribution")
```

This distribution shows a left skew with a slight tail on the lower end of the distribution, from around 0.89 and below. Most of the values are concentrated between 0.90 and 0.94.


Log(beta_1 * beta_2) distribution

```{r}
weather_boot_results %>% 
  ggplot(aes(x = log_b1b2)) + 
  geom_density() +
  labs(title = "log(beta_1 * beta2) distribution")
```

This distribution is heavily left-skewed, with a long tail from about -13 to around -6.5, when it evens out more. The most values are concentrated around -6 to -5.


### 95% confidence intervals for both estimates

```{r}
weather_boot_results %>% 
  pivot_longer(
    cols = c(r_squared, log_b1b2),
    names_to = "estimate_type",
    values_to = "estimate"
  ) %>% 
  na.omit() %>% 
  group_by(estimate_type) %>% 
  summarize(
    ci_lower = quantile(estimate, 0.025),
    ci_upper = quantile(estimate, 0.975)
  )
```


The 95% CI for the r_squared estimates is (0.889, 0.940). The 95% CI for the log(beta_1 * beta_2) is (-9.43, -4.52). This aligns with the plots.










